---
title: "meeting 1"
author: "Blain Morin"
date: "September 9, 2018"
output: 
  html_document:
    theme: journal
---

```{r, echo=FALSE, results='hide', message=FALSE, warning=FALSE}
require(dplyr) || install.packages("dplyr")
require(meta) || install.packages("meta")
require(metasens) || install.packages("metasens")
require(readr) || install.packages("readr")
require(knitr) || install.packages("knitr")
require(kableExtra) || install.packages("kableExtra")
```


Here is the structure of the data from Orestis (no_of_studies > 5).

```{r, echo = FALSE, message=FALSE, warning=FALSE}

X2010_6 <- read_delim("2010_6.csv", "\t", 
                      escape_double = FALSE, trim_ws = TRUE)

small.data = X2010_6 %>% 
          filter(number_review >= 1, 
                 number_review <= 500,
                 no_of_studies >= 5)

### Add number of obs column 
small.data = small.data %>% mutate(obs = a + b + c +d)

kable(head(small.data)[1:5,]) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))
```

All of the column names:
```{r, echo=FALSE}
colnames(small.data)
```


Using meta package to calulate odds ratios for each outcome.
```{r}
### Function to retrieve odds ratio

get.or = function (a, b, c, d){
  
  the.meta = metabin(event.e = a, n.e = a + c, event.c = b, n.c = b + d, sm = "OR")
  exp(the.meta$TE.fixed)
  
}

### Function to retrieve CI lower bound

get.lb = function (a, b, c, d){
  
  the.meta = metabin(event.e = a, n.e = a + c, event.c = b, n.c = b + d, sm = "OR")
  exp(the.meta$lower.fixed)
  
}

### Function to retrieve CI upper bound

get.ub = function (a, b, c, d){
  
  the.meta = metabin(event.e = a, n.e = a + c, event.c = b, n.c = b + d, sm = "OR")
  exp(the.meta$upper.fixed)
  
}
```

Here I retrieve the odds ratios and CIs for each outcome:
(this would be the "truth"")

```{r, cache=TRUE}

truth = small.data %>% group_by(my_ID) %>%
  mutate(truth.OR = get.or(a = a, b = b, c = c, d = d),
         lowerCI = get.lb(a = a, b = b, c = c, d = d),
         upperCI = get.ub(a = a, b = b, c = c, d = d))

```

```{r, echo = FALSE}

truth.table = truth %>% 
  group_by(my_ID) %>%
  slice(1)

kable(truth.table[1:6, c(1:2, 9,11, 21:23)], digits = 3)


```


Example function for dropping a proportion of studies because of small sample size:

```{r}

drop.proportion = function(df, proportion) {
  
  new.df = df %>%
    group_by(.data$my_ID) %>%
    arrange(desc(.data$obs), .by_group = TRUE) %>%
    slice( seq((1-proportion) * n()) )
  
  return(new.df)
  
}

```

Example: Remove the smallest 20% (by sample size)

```{r}

biased.data = drop.proportion(small.data, .2)

### Recalculate OR and CI

half.truth = biased.data %>% group_by(my_ID) %>%
  mutate(OR = get.or(a = a, b = b, c = c, d = d),
         lowerCI = get.lb(a = a, b = b, c = c, d = d),
         upperCI = get.ub(a = a, b = b, c = c, d = d))

```

```{r, echo = FALSE}
half.truth.table = half.truth %>% 
  group_by(my_ID) %>%
  slice(1)

kable(half.truth.table[1:6, c(1:2, 9,11, 22:24)], digits = 3)
```

Trim and fill on the biased set (using meta package):

```{r, message=FALSE}

added.trim = half.truth %>%
  group_by(my_ID) %>% 
  mutate(new_n_of_studies = n()) %>%
  filter(new_n_of_studies >= 5) %>%
  mutate(trimfill.OR = exp(trimfill( metabin(event.e = a, n.e = a + c, event.c = b, n.c = b + d, sm = "OR"))$TE.random))

interm = truth %>%
  select(-lowerCI, -upperCI)

joinedwithtruth = inner_join(x = added.trim, y = interm) %>%
  group_by(my_ID) %>%
  slice(1)

```

```{r, echo=FALSE, message=FALSE}

ratio.table = joinedwithtruth %>%
  select(review, comparison, OR, trimfill.OR, truth.OR)

kable(ratio.table[1:10,], digits = 3)

```